### 2016-04-26 paper/proposal.tex
Added latex sources.

\author{Group 40: Arnold, Gurunat, Huang, \& Qu}
\title{Modular Deep Encoder-Decoders}
\date{\today}
\documentclass[12pt]{article}

\input{\string~/tex_templates/core}

% Custom Title

%\def\maketitle{
    %\centering
    %\par\textbf{\LARGE\@title}
    %\par\hfill
    %\par{\@author, \@date}
    %\par\hfill
    %\par\hfill
    %\rule{\textwidth}{3pt}
%}

\def\maketitle{
    \begin{centering}
    \par\rule{\textwidth}{2pt}
    \par\hfill
    \par\textbf{\LARGE\@title}
    \par\hfill
    \par{\textit{\@author}}
    \par\hfill
    \par{\@date}
    \par\rule{\textwidth}{2pt}
    \end{centering}
}

\begin{document}
\thispagestyle{empty}
\maketitle
\hfill
\begin{abstract}
In this short paper we propose an approach to transfer learning using rich
vector embeddings. The suggested technique can be applied to any supervised
task, and it handles multiple sources and changing sources of data without the
need for retraining. To verify our ideas, we apply our ideas to the task of
text-classification.
\end{abstract}

\section{Introduction}\label{introduction}

Our goal is to generate rich vector embeddings from articles to classify
them into predefined categories. Then, we want to extend our model with
an additional source of data: the title of the videos. To handle this
new knowledge source without retraining our previous model, we suggest
to generate a new embedding that will be used to modify the original
one. This combined embedding will then be used for the classification
task.

More formally, we want to jointly learn a set of \emph{encoder}
functions $\{E_i\}_0^N$ mapping samples $x \sim \chi_i$ from a set of
data distributions $\{\chi_i\}_0^N$ to a fixed-sized vector embedding
$V$.

\[ \forall i \in [0; N]: E_i(x): \chi_i \rightarrow V_i \in \Re^M\]

\[V = \sum_i^N V_i\]

The embedding $V$ is then fed into a decoder function $D$ which in the
case of classification learns a mapping from the vector space of $V$ to
the label space $L$.

\[ D(v): \Re^M \rightarrow L \in \Re^D\]

Finally, we want to extend the set of encoders by learning a new encoder
$E_{N+1}$ which handles samples from a different dataset $\chi_{N+1}$,
without retraining our trained decoders and encoders.

\[ E_{N+1}(x): \chi_{N+1} \rightarrow V_{N+1} \in \Re^M\]

Although we could have used any kind of mapping, we chose to use deep
learning algorithms, as they easily learn hierarchical representations
and have been known to highly outperform other statistical techniques on
natural language tasks. Learning embeddings has been previously done by
\href{References}{{[}1{]}}, although our proposed contribution is
slightly different. \href{References}{{[}1{]}} used deep autoencoders to
obtain a better initialization of the parameters of their model. Our
approach instead is much closer in spirit to the work of Sutskever
\href{References}{{[}3{]}} and Vinyals \href{References}{{[}4{]}}. They
both use encoders on a sequence of data to generate a
\emph{thought-vector} which will then be decoded in the desired terget
sequence. In some sense, our proposal adds the transfer learning
component to their contribution. This approach is also similar to the
work of Karpathy \& al \href{References}{{[}5{]}} where they mapped
images to their captions with embeddings.

\section{Method}\label{method}

\subsection{Materials}\label{materials}

We will use the freely avaiable dataset from
\url{https://github.com/ParallelMazen/SaudiNewsNet} and extract the
titles and articles content. Since the dataset initial purpose is not
the classification, we will have to slightly rework it to fit our
purpose.

We will use the neon python library for training deep networks. In the
case where the training phase would become too expensive, we will also
rely on the mpi4py library.

\subsection{Procedure}\label{procedure}

The procedure is quite straight-forward. While part of the team will
work on building our tailored dataset, the other half will work on the
model definition and training procedure. Coordination at the beginning
of the work will be curcial. In particular, we plan on feeding the text
data as a sequence of one-hot character encodings.

Once every pre-requisite is available, we will train the first encoder
$E_1$ (implemented as a recurrent neural network) to build the embedding
$V_1$. Since we will only be dealing with a relatively simple
classification task, our decoder $D_1$ will simply be a fully connected
multi-layer perceptron. They will be jointly trained end-to-end by
propagating the gradients through the embedding from $D_1$ to $E_1$.

The second training step will be to train the second encoder $E_2$.
Again, we also perform training end-to-end, but specifically do not
propagate the gradients through $E_1$. We hope to avoid optimizing the
parameters of $D_1$, but this might not be practically doable.

\subsection{Evaluation}\label{evaluation}

The final evaluation scheme is mostly to be determined depending on our
obtained outcomes. Obviously, we want to compare the new model with the
the the jointly trained one, as well as a classification baseline.

\section{Participants \& Labor
Division}\label{participants-labor-division}

The participants in alphabetical order:

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Seb Arnold - 9013085897 - arnolds@usc.edu
\item
  Prashanth Gurunath Shivakumar - 2251924199 - pgurunat@usc.edu
\item
  Qiangui Huang - 9532576000 - qianguih@usc.edu
\item
  Cheng Qu - 2385279985 - chengqu@usc.edu
\end{itemize}

Instead of having a fixed division of the labor, we assigned one
responsible to each task. The responsible acts as a project lead on the
designed task and other team members can contrbibute to the task as
required.

\begin{itemize}
\item
  Dataset: Cheng
\item
  Model Training: Quiangui
\item
  Knowledge Transfer: Seb
\item
  Report Writing: Prashanth
\end{itemize}

\section{References}\label{references}

Part of the relevant literature review. More literature was involved for
the deep learning part.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Supervised Representation Learning: Transfer Learning with Deep
  Autoencoders, Zhuang \& al,
  http://ijcai.org/Proceedings/15/Papers/578.pdf
\item
  Transfer Learning via Dimensionality Reduction, Pan \& al,

  https://www.cse.ust.hk/\textasciitilde{}jamesk/papers/aaai08.pdf
\item
  Sequence to Sequence Learning with Neural Networks, Sutskever \& al,

  http://arxiv.org/abs/1409.3215
\item
  Grammar as a Foreign Language, Vinyals \& al,
  http://arxiv.org/abs/1412.7449
\item
  Deep Fragment Embeddings for Bidirectional Image Sentence Mapping,
  Karpathy \& al, https://cs.stanford.edu/people/karpathy/nips2014.pdf
\end{enumerate}

%\end{multicols}
\end{document}